---
title: 'Cleaning Business: Attribute Objects (Part 3)'
author: "Jonathan Chang"
date: "November 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
library(dplyr)
```

## Objects

Objects in Python, Javascript, etc., might as well be a list of key-value pairs, where the values can be any variable, or even functions (since functions are first class citizens). Objects are often passed as JSON (Javascript Object Notation), which are string-string key-value pairs of a specific format. Since the Yelp data set was at some point obtained from its REST API, it chose to use objects to attributes with several related properties.

The problem, again, is that machine learning algorithms will recognize these as mere text or factors -- equally useless, when they're expecting numbers. Treating combinations of multiple properties as text also fails to preserve the relations that they hold. So, again, we'll split these into separate columns.

To make our lives easier, we'll leverage existing libraries that split JSON into separate variables. We'll make a few changes to make these attribute objects compatible with JSON format.

```{r, warning=FALSE}
library(jsonlite)
convert_json = function(attr) {
  txt = gsub("'", "\"", attr)
  txt = gsub("False", "false", txt)
  gsub("True", "true", txt)
}
```

## Splitting

```{r}
business = read.csv('business_preclean2.csv', stringsAsFactors=FALSE)
```

As before, we'll split these into separate columns. Let's see an example before we start.

```{r, echo=FALSE}
business$attributes_Music[55]
```

```{r}
# Inputs
# data - dataframe
# attr - attribute name without the "attributes_" prefix
expand_attr = function(data, attr) {
  
  # To keep new columns organized, we'll prepend the original
  # attribute name to each new column.
  prefix = attr
  
  # Add "attributes_" back in the column title, so we don't
  # have to type this over and over.
  attr = paste("attributes", attr, sep='_')

  # Convert column to JSON format.
  txt = convert_json(data[,attr])
  df = data.frame()
  
  for (row in 1:length(txt)) {
    
    # Create a list using number of columns of new dataframe. 
    cols = ncol(df)
    df[row,] = rep("None", each=cols)
    
    # Attribute is not empty.
    if (nchar(txt[row]) > 0) {
      
      # Extract properties from JSON attribute object.
      js = fromJSON(txt[row])
      colns = colnames(df)
      
      # Assign value of each property 
      # (add column if we haven't seen it before).
      for (category in 1:length(js)) {
        cat_value = js[category]
        cat_name = trimws(names(cat_value))
        if (!(cat_name %in% colns)) {
          df[cat_name] = "None"
        }
        df[row, cat_name] = ifelse(cat_value, 1, 0)
      }
    }
  }
  
  # Add original attribute name as prefix to all new columns.
  colns = colnames(df)
  colns = gsub('-', '', colns)
  colns = as.vector(sapply(colns, function(x) paste(prefix, x, sep='_')))
  colnames(df) = colns
  
  # Replace original attribute column with new columns.
  data = cbind(df, data)
  data[,attr] = NULL
  data
}

# Apply to every attribute that uses objects.
business = expand_attr(business, "Music")
business = expand_attr(business, "HairSpecializesIn")
business = expand_attr(business, "GoodForMeal")
business = expand_attr(business, "DietaryRestrictions")
business = expand_attr(business, "BusinessParking")
business = expand_attr(business, "BestNights")
business = expand_attr(business, "Ambience")
```

We now have `r ncol(business)` columns. But that's the cost of doing business. The next best alternative is to delete these attributes altogether rather than keeping them in a useless format. Let's see the next columns before we save.

```{r}
business[55, grepl("Music_", names(business))]
```

Finally, let's change all blank values into "None" to keep it consistent.

```{r}
business[business==""] = "None"
business[is.na(business)] = "None"
```

```{r, eval=FALSE}
write.csv(business, 'business_preclean3.csv', row.names=FALSE)
```